---
title: "Practical Machine Learning: \n Identifying bad tecnhique in lifting"
author: "Sergio"
date: "8/4/2021"
output: html_document
---

Machine learning is very nice. This is the final project from the course Practical Machine Learning, given by the John Hopkins University via Coursera.

In this project, the goal was be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. So, the idea . The data was kindly delivered by the [Human Activity Recognition](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har) Project.

Both the  [training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and the [Final Test](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv) are available for the public.

# Loading the required libraries

```{r setup}
library(caret)
library(ggplot2)
library(randomForest)
```

# Reading and cleaning the data

```{r read, cache= T}
Training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
Final_Test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
dim(Training)
dim(Final_Test)
```

Data for training has 19622 observations on 160 variables, and the testing data set has only 20 observations with 160 variables.


#  Preparing a tidy data set

Let's see what the data set is
```{r, cache= T}
str(Training)
```

So, first variable is just an index and I don't want it. neither do I want the username cause I don't consider it a useful variable. Similar with all the 7 first variables
```{r, cache= T}
Training <- Training[,-c(1:7)]
dim(Training)
```

There is also a lot of variables with a bunch of NA's and I don't want them.
```{r, cache= T}
mynearzero <- nearZeroVar(Training)
Training <- Training[,-mynearzero]
dim(Training)
```

Also, there is a bunch of columns with missing data. So, in the next line, I will erase all those columns whose NA values are more than 60% of the number of rows.
```{r, cache= T}
Training <- Training[, colSums(is.na(Training)) <= nrow(Training)*0.6]


```

Finally, let's convert the data columns to the right format. All are numeric except the last, which is class, the one we actually want to predict.
```{r, cache= T}
cols <- c(1:52)
Training[,cols] <- sapply(Training[,cols],as.numeric)
Training$classe <- as.factor(Training$classe)
```

# Splitting the training data set

The Final Test data set will only be used by the course to demonstrate that the model is working, and it is very small. Thus, I have to split my "training" data set into two datasets, one will actually work to train the model and the other to test it before submission. I will choose 60% of the data to train and 40% to test. This is kindly done by the package caret.

```{r split, cache= T}
toTrain <- createDataPartition(y=Training$classe, p=0.6, list=FALSE)
TrainSet <- Training[toTrain, ]
TestSet <- Training[-toTrain, ]
dim(TrainSet); dim(TestSet)
```

# Building a model

There is a ton of different models I could try. I am going to start with Random Forests because we know it's highly accurate. We know it's less scalable, but for the moment that is not important.

```{r model, cachet= T}
#Creating the model
RF_model <- randomForest(classe~. , data=TrainSet)

#Using it to predict
RF_predictions <- predict(RF_model, TestSet, type = "class")

#Checking how predictions went with my mini test set.
confusionMatrix(RF_predictions, TestSet$classe)
```
So, accuracy is 99,45%. High enough, I will take it. I could try other models and test them, and maybe combine them. But as I am typing I also have a pizza in the oven, so let's leave it like this.


# A final challenge with 20 incognito data points 

Only Coursera and God will know if it's wrong or if it's right...

```{r testing}
mYpRedIctIonS <- predict(RF_model,Final_Test)
mYpRedIctIonS

```
